The impact of forest management on the temperature sensitivity of organic matter turnover

### 2024-11-14 Manuscript_draft/Draft.Rmd
small changes in discussion

## Role of edaphic, ecological and microbiological variables in temperature sensitivity
The Q10 values obtained in our study were comparable to the Q10 values measured in the field by other studies (ref). Xu et al 2015 observed the largest relative frequency of Q10 values across forest in China was within the range of 2.0–3.0. Fierer et al. (2006) observed Q10 values ranging from 2.2 to 4.6 (average of 3.0) at continental-scale in an incubation experiment. Still, we observed a large variation of Q10 depending on the forest, being the highest in Romania. This is in agreement with the known higher temperature sensitivity in colder regions (Koven et al 2017). The reason could be related to longer time of frozen soils that could reduce the diffusion of substrates and extracellular enzymes within the soil (Davinson and Jansen 2006), giving higher values of Q10 not related to recalcitrance.

Our analyses revealed the three variables that explained most of the variation of Q10 were related to i) the quality of substrate (i.e., nitrogen content as limiting nutrient in forest), ii) the microorganisms, in particular, fungal to bateria ratio and iii) the primary productivity (that is, the number of trees).

>>>>

## Role of edaphic, ecological and microbiological variables in temperature sensitivity
The Q10 values obtained in our study were comparable to the Q10 values measured by other studies. Xu et al 2015 observed the largest relative frequency of Q10 values across forest in China was within the range of 2.0–3.0. Fierer et al. (2006) observed Q10 values ranging from 2.2 to 4.6 (average of 3.0) at continental-scale in an incubation experiment. Still, we observed a large variation of Q10 depending on the forest, being the highest in Romania. This is in agreement with the known higher temperature sensitivity in colder regions (Koven et al 2017). Longer time of frozen soils in Romania could have reduced the diffusion of substrates and extracellular enzymes within the soil (Davinson and Jansen 2006), giving higher values of Q10 not related to recalcitrance.

Our analyses revealed the three variables that explained most of the variation of Q10 were related to i) the quality of substrate (i.e., nitrogen content as limiting nutrient in forest), ii) the microorganisms, in particular, fungal to bacteria ratio and iii) the soil temperature. Our results indicate that SOM decomposition not only depends on the substrate biochemistry but also on the ability of the existing microbial community to decompose the available substrate. This is in partial agreement with a recent global scale study stating that soil microbiomes are the main players governing the responses of soil heterotrophic respiration to temperature, and secondly substrate quantity and biochemical recalcitrance (Sáez-Sandino et al 2023). Since total microbial biomass did not correlated with total nitrogen, the observed effect of nitrogen might be related to an increased in mineral nitrogen.
Compared to controls, fungal and bacteria biomass decreased in both thinning and clear-cut indicating a stronger decrease in ectomycorrhizal fungi compared to root-fungi associated bacteria. In forest, the role of fungi, especially of ectomycorrhizal fungi in decomposing SOM while mining for nutrients. The stronger decrease in ectomycorrhizal fungi, might have increased sensitivity to temperature linked to reduced fungal activity. The decrease This could   The higher increase in fungal biomass compared to bacteria in the treatments with slash indicate a faster colonization of saprotrophic fungi in litter. An probably temperatre sensitivity is controlled by a reduced number of taxa with a high abundance (Sáez-Sandino et al 2023). Diversity metrics would give more insights about that.
